# Hate Crime Data {#hate_crimes}

```{r, echo=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  error   = FALSE,
  message = FALSE
)
```

```{r}
load("data/ucr_hate_crimes_1991_2018.rda")
```


This dataset covers crimes that are reported to the police and judged by the police to be motivated by hate More specifically, they are, first, crimes which were, second, motivated - at least in part - by bias towards a certain person or group of people because of characteristics about them such as race, sexual orientation, or religion. The first part is key, they must be crimes - and really must be the selection of crimes that the FBI collects for this dataset. Biased actions that don't meet the standard of a crime, or are of a crime not included in this data, are not considered hate crimes. For example, if someone yells at a Black person and uses racial slurs against them, it is clearly a racist action. For it to be included in this data, however, it would have to extend to a threat since "intimidation" is a crime included in this data but lesser actions such as simply insulting someone is not included. For the second part, the bias motivation, it must be against a group that the FBI includes in this data. When this data collection began crimes against transgender people were not counted so if a transgender person was assaulted or killed because they were transgender, this is not a hate crime recorded in the data.^[The first year where transgender as a group was a considered a bias motivation was in 2013.] 

So this data is really a narrower measure of hate crimes than it might seem. In practice it is (some) crimes motivated by (some) kinds of hate that are reported to the police. It is also the most under-reported UCR dataset with most agencies not reporting any hate crimes to the FBI. This leads to huge gaps in the data with some states having zero agencies report crime, agencies reporting some bias motivations but not others, agencies reporting some years but not others. While these problems exist for all of the UCR datasets, it is most severe in this data. This problem is exacerbated by hate crimes being rare even in agencies that report them - with such rare events, even minor changes in which agencies report or which types of offenses they include can have large effects. 

My main takeaway for this data is that it is inappropriate to use for most hate crime research. At most it can be used to look at within-city within-bias-motivation trends, while keeping in mind that even this narrow subset of data is limited by under-reporting by victims and potential changes in police practices of reporting such as how many months of data they report per year. 


## Agencies reporting

- hate crime data is available from 1991 to 2019 - 

```{r hateAgencies, fig.cap = "The annual number of police agencies that report at least one hate crime incident in that year."}
ucr_hate_crimes_1991_2018 %>%
  dplyr::filter(hate_crime_incident_present_flag %in% "one or more hate crime incidents present") %>%
  dplyr::distinct(ori, year, .keep_all = TRUE) %>%
  count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line(size = 1.05) +
  xlab("Year") +
  ylab("# of Agencies") +
  theme_crim() +
  scale_y_continuous(labels = scales::comma) +
  expand_limits(y = 0)
```

Another way to understand reporting is to look at the number of reported hate crimes by state and see which states report and which don't. Figure \@ref(fig:hateCrimesMap) does this for 2018 data by showing the number of reported hate crime incidents by state. While every state other than Wyoming reporting at least one hate crime in 2018, there are large differences between states because even in states that have reporting agencies, not all agencies in that state report. For example, Alabama reported only two hate crimes in 2018, the least of any state other than Wyoming. 

```{r, hateCrimesMap, fig.cap="Total reported hate crimes by state, 2018."}
ucr_hate_crimes_1991_2018 %>%
  dplyr::filter(year %in% 2018,
                !is.na(incident_number),
                !is.na(state)) %>%
  dplyr::count(state) %>%
  ggplot2::ggplot(aes(map_id = state)) + 
  ggplot2::geom_map(aes(fill = n), map = fifty_states, color = "black") + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "", fill = "Hate Crimes") +
  theme(panel.background = element_blank()) +
  fifty_states_inset_boxes() +
  scale_fill_gradient(low = "white", high = "red", breaks=c(50, 250, 400, 550, 700, 850, 1000)) 
```

Since the number of state-wide hate crimes is partially influenced by population, we'll also look at it as the percent of agencies in the state that report at least one hate crime. Again this is limited by population as agencies in each state cover different populations - and most agencies are small so a few large agencies can account for a large share of crimes, including hate crimes. Still, this is a useful exercise as we can easily see where very few agencies report. Figure \@ref(fig:statePercentReporting) shows the percent of agencies for each state that reported at least one hate crime in 2018. In Connecticut, the state with the highest percent of agencies reporting, 28% of agencies reported at least one hate crime. In 34 states, fewer than 10% of agencies reported a hate crime, and in 19 states fewer than 5% of agencies did so. One interesting finding from this graph is the more liberal states - Connecticut, New Jersey, Vermont, California, Washington - have the highest share of agencies reporting a hate crime, indicating that the culture of the state may influence whether agencies report hate crimes.   

```{r statePercentReporting, fig.cap = "The percent of agencies in each state that reported at least one hate crime in 2018."}
ucr_hate_crimes_1991_2018$hate_crime_binary <- 0
ucr_hate_crimes_1991_2018$hate_crime_binary[ucr_hate_crimes_1991_2018$hate_crime_incident_present_flag %in% "one or more hate crime incidents present"] <- 1
ucr_hate_crimes_1991_2018$state_abb_ori <- substr(ucr_hate_crimes_1991_2018$ori9, 1, 2)
temp <- ucr_hate_crimes_1991_2018 %>%
  dplyr::filter(year %in% 2018,
                !is.na(state_abb_ori)) %>%
  dplyr::distinct(ori9, .keep_all = TRUE) %>%
  dplyr::group_by(state_abb_ori) %>%
  dplyr::summarize(prop_reporting = mean(hate_crime_binary)) %>%
  dplyr::mutate(percent_reporting = prop_reporting * 100)
state <- datasets::state.name[match(tolower(temp$state_abb_ori), tolower(datasets::state.abb))]
temp$state <- state
temp <- temp[!is.na(temp$state), ]
temp$state <- tolower(temp$state)

ggplot2::ggplot(temp, aes(map_id = state)) + 
  ggplot2::geom_map(aes(fill = percent_reporting), map = fifty_states, color = "black") + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "", fill = "% of Agencies With a Hate Crime") +
  theme(panel.background = element_blank()) +
  fifty_states_inset_boxes() +
  scale_fill_gradient(low = "white", high = "red", breaks=c(1, 5, 10, 15, 20, 25, 30)) 

```

## Important variables

Each hate crime incident can cover up to 10 different crimes occurring - e.g. a person who burglarizes a synagogue and spray paints a swastika on the wall would be have both vandalism and burglary reported in this data. With each crime, this data has the bias motivation for that crime, the location of the crime (in broad categories, not the actual location in the city like a street address would have), and the number of victims for that offense. In practice, in hate crimes with multiple offenses recorded, the bias motivation, location, and victim count is the same for each offense. 




### Bias motivation

The most important variable in this data is the "bias motivation" which is the FBI's term for the cause of the hate. A hate crime targeted against Black people would be an "anti-Black" bias motivation. For the police to classify an incident as a hate crime, and to assign a particular bias motivation, the police must have *some* evidence that the crime was motivated by hate. The victim saying that the crime is a hate crime alone is not sufficient - though if large portions of the victim's community believe that the crime is a hate crime, this is a factor in the police's assessment. The evidence required is not major, it includes things as explicit as racial slurs said during an incident and less obvious factors like the victim is celebrating their community (e.g. attending a holiday event) or the crime occurring on an important holiday for that community (e.g. Martin Luther King Day, religious holidays). The FBI also encourages police to consider the totality of the evidence even if none alone strongly suggests that the crime was a hate crime in making their determination about whether the incident was a hate crime or not.

This also means that many (likely most) hate crimes will not be recorded as hate crimes since there is no evidence that the crime is motivated by hate. For example, if a man targeted Asian people for crimes because they are Asian, that would in reality be a hate crime. However, if the offender does not say anything anti-Asian to the victim, which is the mostly likely thing to indicate that this is a hate crime, the crime would not likely be recorded as a hate crime. 

Indeed, even if crimes against Asian people significantly increase, without evidence 

Bias motivation is based on the offender's perceptions of the victim so even if they are incorrect in who their victim is, if they intended to target someone for their perceived group membership, that is still a hate crime. For example, if a person assaults a man because they think he is gay, that is a hate crime because the assault was motivated by hate towards gay people. Whether the victim is actually gay or not is not relevant - the offender perceived him to be gay so it is an anti-gay  hate crime. To make this even more complicated, the offender must have committed the crime because they are motivated, at least to some degree, by their bias against the victim. Being biased against the victim but targeting them for some other reason means that the crime is not a hate crime. 



```{r hateBiasMotivation, fig.cap = "The bias motivation for hate crime incidents. In incidents with multiple bias motivation, this shows only the first bias motivation recorded."}
ucr_hate_crimes_1991_2018$bias_motivation_offense_1[ucr_hate_crimes_1991_2018$bias_motivation_offense_1 %in% "99"] <- NA
temp <- make_frequency_table_year(ucr_hate_crimes_1991_2018, "bias_motivation_offense_1",
                             c("Bias Motivation", "First Year Reported", "# of Incidents", "% of Incidents"))
temp$`Bias Motivation` <- gsub("lgbt", "LGBT", ignore.case = TRUE, temp$`Bias Motivation`)
datatable(temp, options = list(pageLength = 100)) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```
### Which crime occurred

```{r hateOffense, fig.cap = "The offense type for hate crime incidents. In incidents with multiple offense types, this shows only the first offense type recorded."}
ucr_hate_crimes_1991_2018$ucr_offense_code_1[ucr_hate_crimes_1991_2018$ucr_offense_code_1 %in% "720"] <- NA
temp <- make_frequency_table_year(ucr_hate_crimes_1991_2018, "ucr_offense_code_1",
                             c("Offense", "First Year Reported", "# of Incidents", "% of Incidents"))
datatable(temp, options = list(pageLength = 100)) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```
### Location

```{r hateLocations, fig.cap = "The location of hate crime incidents. In incidents with multiple locations, this shows only the first location recorded."}
ucr_hate_crimes_1991_2018$location_code_offense_1[ucr_hate_crimes_1991_2018$location_code_offense_1 %in% "undocumented code v13"] <- NA
temp <- make_frequency_table_year(ucr_hate_crimes_1991_2018, "location_code_offense_1",
                             c("Location", "First Year Reported", "# of Incidents", "% of Incidents"))
datatable(temp, options = list(pageLength = 100)) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```
### Number and race of offenders

There are two variables that have information about the people who commit the hate crime: the number of offenders and the race of the offenders (as a single value with the race of the group if all are of the same race or it will say a "multi-racial" group). Unfortunately, important information like the age of the offenders, their criminal history, their relationship to the victim, their gender, or whether they are arrested are completely unavailable in this dataset. 

When the police do not have any information about the number of offenders (which is common in cases of property crimes such as vandalism but rare in violent crimes), this data considers that to have zero offenders. The zero is just a placeholder that means that the police have no idea how many offenders there are, not that they think there were actually no offenders. Figure \@ref(fig:hateCrimeOffenderNumber) shows the percent of hate crimes from 1991-2018 that have each number of offenders recorded. In the actual data it says the actual number of offenders, with the largest group in the current data going to 99 offenders - in this graph I group 10 or more offenders together for simplicity. I also relabel zero offenders as "Unknown" offenders in the graph since that's more accurate. The most common number of offenders per hate crime is one offender, at about 46% of hate crimes from 1991-2018 having only one offender. This drops sharply to 9% of hate crimes having 2 offenders and continues to drop as the number of offenders increase. However, about a third (36.8%) of hate crimes have an unknown number of offenders. 

```{r hateCrimeOffenderNumber, fig.cap = "The race of offenders, as a group, for hate crime incidents, 1991-2018."}
temp <- ucr_hate_crimes_1991_2018 %>%
  filter(hate_crime_incident_present_flag %in% "one or more hate crime incidents present") 
temp$total_offenders[temp$total_offenders >= 10] <- "10 or more"
temp$total_offenders[temp$total_offenders %in% 0] <- "Unknown"
temp$total_offenders <- factor(temp$total_offenders, levels = rev(c(1:9, "10 or more", "Unknown")))
  ggplot2::ggplot(temp, aes(x = total_offenders)) + 
  theme_crim() + 
  ggplot2::coord_flip() +
  ggplot2::xlab("") +
  ggplot2::geom_bar(ggplot2::aes_string(y = "(..count..)/sum(..count..)")) + 
  crimeutils::theme_crim() +
  ggplot2::scale_y_continuous(labels = scales::percent, 
                              expand = c(0, 0)) +
  ylab("% of Hate Crimes")
```


The data also includes the race of the offenders as a group, though not the ethnicity (Hispanic or not) of the offenders. In cases where there are multiple races among offenders, that will be classified as a "multi-racial group". As shown in Figure \@ref(fig:hateCrimeOffenderNumber) The most common racial group is "unknown" since the police do not know the race of the offenders. Then are White offenders at nearly 40% of hate crimes followed by Black offenders at nearly 13% of hate crimes. The remaining racial groups are rare with about 2% of hate crimes being committed by a multi-racial group of offenders and 0.72% of hate crimes committed by Asian or Pacific Islander offenders and 0.54% committed by American Indian or Native Alaskan offenders. 

```{r hateCrimeOffenderRace, fig.cap = "The race of offenders, as a group, for hate crime incidents, 1991-2018."}
ucr_hate_crimes_1991_2018 %>%
  filter(hate_crime_incident_present_flag %in% "one or more hate crime incidents present") %>%
  tidyr::replace_na(list(offenders_race_as_a_group = "unknown")) %>%
  dplyr::mutate(offenders_race_as_a_group = crimeutils::capitalize_words(offenders_race_as_a_group)) %>%
crimeutils::make_barplots(column = "offenders_race_as_a_group", count = FALSE) +
  ggplot2::scale_y_continuous(labels = scales::percent, 
                expand = c(0, 0)) +
  ylab("% of Hate Crimes")
```
