# Supplementary Homicide Reports (SHR) {#shr}

```{r, echo=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  error   = FALSE,
  message = FALSE
)
```

```{r}
library(crimeutils)
library(dplyr)
library(readr)
library(DT)
library(scales)
library(ggplot2)
load("data/shr_1976_2018.rda")



make_frequency_table <- function(data, column, col_names) {
  temp <- unique(data[, column])
  temp <- temp[!is.na(temp)]
  temp_df <- data.frame(col1 = temp,
                        number   = NA) 
  
  for (i in 1:nrow(temp_df)) {
    loop_value <- temp_df$col1[i]
    storage <- data[data[, column] %in% loop_value, ]
    temp_df$number[i] <- nrow(storage)
  }
  temp_df <-
    temp_df %>%
    mutate(percent = number / sum(number)) %>%
    arrange(desc(number)) %>%
    mutate(col1 = crimeutils::capitalize_words(col1))
  names(temp_df) <- col_names
  return(temp_df)
}


```

```{r shrOffenderAge, fig.cap = "The age of homicide offenders, based on the first offender in any homicide incident. Victims under age 1 (classified as 'birth to 7 days old, including abandoned infant' and '7 days to 364 days old') and considered 0 years old. Victims reported as '99 years or older' are considered 99 years old."}
temp <- shr_1976_2018
temp$offender_1_age[temp$offender_1_age %in% c("bb", "nb")] <- 0
temp$offender_1_age[temp$offender_1_age %in% "99 years or older"] <- 99
temp$offender_1_age[temp$offender_1_age %in% "unknown"] <- NA
temp$offender_1_age <- readr::parse_number(temp$offender_1_age)
mean(temp$offender_1_age, na.rm = TRUE) #30.96 years old
crimeutils::make_stat_count_plots(temp, 
                                  column = "offender_1_age",
                                  ylab = "# of Offenders",
                                  xlab = "Age",
                                  count = FALSE) +
  geom_vline(aes(xintercept = mean(temp$offender_1_age, na.rm = TRUE)), color = "#d95f02", size = 1.07)
```


```{r shrVictimAge, fig.cap = "The age of homicide offenders, based on the first offender in any homicide incident. Victims under age 1 (classified as 'birth to 7 days old, including abandoned infant' and '7 days to 364 days old') and considered 0 years old. Victims reported as '99 years or older' are considered 99 years old."}
temp <- shr_1976_2018
temp$victim_1_age[temp$victim_1_age %in% c("7 days to 364 days", "birth to 6 days, including abandoned infant")] <- 0
temp$victim_1_age[temp$victim_1_age %in% "99 years or older"] <- 99
temp$victim_1_age[temp$victim_1_age %in% "unknown"] <- NA
temp$victim_1_age <- readr::parse_number(temp$victim_1_age)
mean(temp$victim_1_age, na.rm = TRUE) #32.87 years old
crimeutils::make_stat_count_plots(temp, column = "victim_1_age", ylab = "# of Victims", xlab = "Age") +
  geom_vline(aes(xintercept = mean(temp$victim_1_age, na.rm = TRUE) ), color = "#d95f02", size = 1.07)
```

The Supplementary Homicide Reports dataset - often abbreviated to SHR -  is the most detailed of the UCR datasets and provides information about the circumstances and participants (victim and offender demographics and relationship status) for homicides.^[If you're familiar with the National Incident-Based Reporting System (NIBRS) data that is replacing UCR, this dataset is the closest UCR data to it, though it is still less detailed than NIBRS data.] For each homicide incident it tells you the age, gender, race, and ethnicity of each victim and offender as well as the relationship between the first victim and each of the offenders (but not the other victims in cases where there are multiple victims). It also tells you the weapon used by each offender and the circumstance of the killing, such as a "lovers triangle" or a gang-related murder. As with other UCR data, it also tells you the agency it occurred in and the month and year when the crime happened. 


```{r shrWeapon, fig.cap = "The weapon used in a homicide incident. In cases where there are multiple offenders, shows only the primary weapon for the first offender."}
temp <- make_frequency_table(shr_1976_2018, "offender_1_weapon",
                             c("Weapon", "# of Incidents", "% of Incidents"))
datatable(temp) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```

```{r shrRelationship, fig.cap = "The relationship between the first victim and the first offender in a homicide incident."}
temp <- make_frequency_table(shr_1976_2018, "victim_1_relation_to_offender_1",
                             c("Relationship", "# of Incidents", "% of Incidents"))
datatable(temp) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```

```{r shrCircumstance, fig.cap = "The circumstance of the homicide for the first offender in a homicide incident."}
temp <- make_frequency_table(shr_1976_2018, "offender_1_circumstance",
                             c("Circumstance", "# of Incidents", "% of Incidents"))
datatable(temp) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```

```{r shrSubCircumstance, fig.cap = "The circumstance for the first offender in a homicide incident in cases where the offender is killed. This includes incidents where the only person who dies in the offender."}
temp <- make_frequency_table(shr_1976_2018, "offender_1_subcircumstance",
                             c("Subcircumstance", "# of Incidents", "% of Incidents"))
datatable(temp) %>%
  formatPercentage('% of Incidents', 2) %>%
  formatRound('# of Incidents', digits = 0)
```



While highly detailed compared to other UCR data, there are a number of limitations for this data. 

Since this data is voluntary to 

This    
If this "most detailed" dataset sounds disappointing - and it is! - 

## A brief history of the data

The data is available from the FBI starting in 1975 though, unlike all later years, this year only has information on a single victim and a single offender. For this reason I only release data starting in 1976 where up to 11 victims and 11 offenders are included. This data has been released every year since and the most recent year available is 2019. 

### Changes in definitions

## What does the data look like?

### Raw data



## What variables are in the data?

### Key variables

## Known issues with the data

## Final thoughts